Preparing your manuscript

### 2023-09-28 paper/paper.qmd
Force myself to write text

    - Be set as a single paragraph.
    - Be less than 250 words for all journals except GRL, for which the limit is 150 words.
    - Not include table or figure mentions.
    - Avoid reference citations unless dependent on or directly related to another paper (e.g., companion, comment, reply, or commentary on another paper(s)). AGU’s Style Guide discusses formatting citations in abstracts.
    - Define all abbreviations.
plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: bibliography.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
---

## Section Heading

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum hendrerit facilisis velit sit amet malesuada. Phasellus ornare nibh augue, maximus sodales ex tristique vitae. Vivamus non sollicitudin orci, aliquam placerat metus. Maecenas volutpat orci felis, vel finibus urna consectetur sed. Integer in dui ac dui mollis imperdiet. Quisque sed dapibus nibh. Aenean non luctus leo. Phasellus luctus mauris id aliquet dictum. Aliquam fermentum semper massa, vel dignissim nibh dictum et. See @Hubbard2021.

In luctus mauris vitae imperdiet luctus. Morbi volutpat ligula ut tortor fermentum, eu ornare felis luctus. Donec semper diam vitae mattis posuere. Suspendisse facilisis purus nisi, sit amet egestas ex tempor ut. Cras tortor nulla, euismod at fermentum vel, dictum vel justo. Aenean commodo interdum diam nec placerat. Nunc vestibulum felis at est tincidunt, at euismod dui vestibulum. Nulla venenatis tortor at auctor iaculis. Donec consectetur neque ut sagittis ornare. Nullam pharetra felis tempor suscipit efficitur. Curabitur nibh ex, euismod at congue hendrerit, egestas id mi. Duis porttitor neque in commodo elementum. Fusce vitae fermentum nisi, euismod viverra augue. Curabitur at mi pretium, accumsan purus nec, tempus turpis.

Donec non semper dui, quis aliquet est. Quisque quis sapien at massa ultricies egestas. Duis consequat ultricies erat, a pulvinar nisl vestibulum id. Sed tristique turpis ligula, et tempor lectus iaculis at. Vivamus commodo sapien ac turpis vestibulum dapibus. Morbi tristique arcu metus, et laoreet nisi varius nec. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Fusce sit amet nisl at mauris suscipit aliquet. Nulla vitae dignissim urna. Suspendisse sit amet arcu vitae magna blandit mattis. Vivamus convallis efficitur pulvinar. Sed cursus elit nulla. Sed porta, arcu a euismod pretium, odio dui lacinia lacus, ac vulputate nulla augue eget ex. Nullam consequat ligula sit amet mattis aliquam. Nulla risus urna, ultrices vel ullamcorper id, ornare viverra nunc.

Nunc in lobortis lacus. Duis maximus urna leo, varius sodales arcu interdum nec. Pellentesque imperdiet dolor in leo eleifend dapibus. Ut dapibus, lectus non viverra gravida, ipsum ex faucibus tellus, quis iaculis risus tellus eget augue. Nullam a viverra est. Cras velit nisi, interdum in lacus at, vehicula mattis elit. Curabitur eu viverra purus. Proin pellentesque, metus vitae congue convallis, lorem metus feugiat mi, sit amet auctor purus ligula bibendum ante. Nam id justo scelerisque, rhoncus lectus in, fermentum libero. Donec tincidunt egestas ex ac eleifend. Cras faucibus ipsum a nunc faucibus fermentum. Integer et maximus lacus. Nam dictum nibh id viverra convallis.

>>>>

    - Be set as a single paragraph.
    - Be less than 250 words for all journals except GRL, for which the limit is 150 words.
    - Not include table or figure mentions.
    - Avoid reference citations unless dependent on or directly related to another paper (e.g., companion, comment, reply, or commentary on another paper(s)). AGU’s Style Guide discusses formatting citations in abstracts.
    - Define all abbreviations.
plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: bibliography.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
---
## Introduction

River runoff is an important component of the global water cycle as it comprises about one third of the precipitation over land areas (Hagemann, 2020). Moreover, accurate runoff forecasting, especially over extended periods, is pivotal for effective water resources management, as highlighted by studies such as Yang et al. (2018), Tan et al. (2018), and Fang et al. (2019).

Over the past decades, models for long-term runoff forecasting have been primarily bifurcated into physically based models and data-based models. While the former attempts to emulate intricate and nonlinear physical hydrological processes, the latter hinges on establishing statistical models that delineate the relationship between large-scale climate patterns and catchment runoff.

Machine Learning (ML) models, such as those employing artificial neural networks, support vector machines, adaptive neuro-fuzzy inference systems, and notably, Long Short-Term Memory (LSTM) neural networks, have gained traction for long-term hydrological forecasting due to their commendable performance (Humphrey et al 2016, Huang et al 2014, Ashrafi et al 2017, Yuan et al 2018, Xu et al 2021).

LSTM networks, an evolution of the classical Recurrent Neural Networks (RNNs), have shown stability and efficacy in sequence-to-sequence predictions, such as using climatic indices for rainfall estimation or long-term hydrological forecasting. However, a limitation of LSTMs is their inability to effectively capture two-dimensional structures, an area where Convolutional Neural Networks (CNNs) excel. Recognizing this, we introduce the ConvLSTM, which integrates the strengths of both LSTM and CNN, to extract spatiotemporal features from precipitation fields for predicting river runoff in the Baltic Sea catchment, summarized by 97 inidivual rivers.

Modeling the Baltic Sea is to a large part the result of the quality of the freshwater input, that is used for the simulation. Meier and Kauker(2003) showed that decadal salinity variations of about 1 $g$ $kg^{-1}$ are caused, inter alia, by annual runoff variations. Further, Meier and Kauker (2003) showed that about 50 \% of the decadal salinity variability can be explained by variations in freshwater input into the Baltic Sea.

This paper delves into the application of deep learning, particularly ConvLSTM, to the challenging task of precipitation nowcasting, a domain yet to fully harness the potential of advanced machine learning techniques. We present ConvLSTM as a novel solution to this spatiotemporal sequence forecasting challenge, highlighting its advantages and potential future applications.

## Methods

### LSTM network

The Long Short-Term Memory (LSTM), a specialized form of Recurrent Neural Networks (RNNs), is specifically tailored for modeling temporal sequences. Its unique design allows it to adeptly handle long-range dependencies, setting it apart from traditional RNNs in terms of accuracy. This prowess in modeling long-range dependencies has been validated in various studies [12, 11, 17, 23]. The cornerstone of LSTM's innovation is its memory cell, $c_t$t, which functions as a repository of state information, also refered to as long-term memory. This cell is accessed, modified, and reset through several self-parameterized gates. As new inputs arrive, their information is accumulated into the cell if the input gate is activated. The forget gate $f_t$ defines the percentage of the previous cell status that is stored $c_{t-1}$​. The decision to propagate the latest cell output, $c_t$, to the final state, $h_t$, is governed by the output gate, $o_t$. A significant advantage of this architecture is the memory cell's ability to retain gradients. This mechanism addresses the vanishing gradient problem, where, as input sequences elongate, the influence of initial stages becomes harder to capture, causing gradients of early input points to approach zero. The LSTM's activation function, inherently recurrent, mirrors the identity function with a consistent derivative of 1.0, ensuring the gradient remains stable throughout backpropagation.

One LSTM cell hence maybe expressed as:

$$
\begin{aligned}
i_t &= \sigma(W_{xi} x_t + W_{hi} h_{t-1} + W_{ci} \circ c_{t-1} + b_i) \\
f_t &= \sigma(W_{xf} x_t + W_{hf} h_{t-1} + W_{cf} \circ c_{t-1} + b_f) \\
c_t &= f_t \circ c_{t-1} + i_t \circ \tanh(W_{xc} x_t + W_{hc} h_{t-1} + b_c) \\
o_t &= \sigma(W_{xo} x_t + W_{ho} h_{t-1} + W_{co} \circ c_t + b_o) \\
h_t &= o_t \circ \tanh(c_t)
\end{aligned}
$$

### ConvLSTM network

The FC-LSTM, while adept in many scenarios, falters in encoding spatial information when handling spatiotemporal data due to its reliance on full connections in both input-to-state and state-to-state transitions. Addressing this limitation, our design ensures that all inputs $X_1, \ldots, X_t$, cell outputs $C_1, \ldots, C_t$, hidden states $H_1, \ldots, H_t$, and gates $i_t, f_t, o_t$ of the ConvLSTM are 3D tensors. The last two dimensions of these tensors represent spatial dimensions, specifically rows and columns. Conceptually, these inputs and states can be visualized as vectors positioned on a spatial grid.

In the ConvLSTM, the future state of a specific cell on this grid is determined by the inputs and past states of its neighboring cells. This spatial consideration is integrated by employing a convolution operator in both state-to-state and input-to-state transitions, as illustrated in Fig. 2. The foundational equations for ConvLSTM are:

$$
\begin{aligned}
i_t &= \sigma(W_{xi} \ast X_t + W_{hi} \ast H_{t-1} + W_{ci} \circ C_{t-1} + b_i) \\
f_t &= \sigma(W_{xf} \ast X_t + W_{hf} \ast H_{t-1} + W_{cf} \circ C_{t-1} + b_f) \\
C_t &= f_t \circ C_{t-1} + i_t \circ \tanh(W_{xc} \ast X_t + W_{hc} \ast H_{t-1} + b_c) \\
o_t &= \sigma(W_{xo} \ast X_t + W_{ho} \ast H_{t-1} + W_{co} \circ C_t + b_o) \\
H_t &= o_t \circ \tanh(C_t)
\end{aligned}
$$

Considering the states as representations of moving entities, a ConvLSTM with a larger transitional kernel can capture rapid movements, while a smaller kernel is apt for slower motions. Drawing parallels with [16], the traditional FC-LSTM's inputs, cell outputs, and hidden states can also be envisioned as 3D tensors, with the last two dimensions being unitary. Thus, the FC-LSTM can be seen as a specific instance of ConvLSTM where all features are concentrated on a singular cell.

To maintain consistency in the dimensions of states and inputs, padding is essential before convolution operations. This padding, especially at boundary points, can be interpreted as leveraging the state of external surroundings for computations. Typically, all LSTM states are initialized to zero before the first input, symbolizing a lack of knowledge about future events. Similarly, zero-padding (as adopted in this study) on hidden states sets the external world's state to zero, implying no prior information about external conditions. This padding approach allows for differential treatment of boundary points, proving advantageous in many scenarios. For instance, in a system with a ball moving and bouncing off unseen walls, the existence of these walls can be inferred by observing the ball's repeated bounces, a deduction challenging if boundary points share the same state transition dynamics as inner points.

### 2023-09-28 paper/paper.tex
Force myself to write text

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{section-heading}{%
\section{Section Heading}\label{section-heading}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum
hendrerit facilisis velit sit amet malesuada. Phasellus ornare nibh
augue, maximus sodales ex tristique vitae. Vivamus non sollicitudin
orci, aliquam placerat metus. Maecenas volutpat orci felis, vel finibus
urna consectetur sed. Integer in dui ac dui mollis imperdiet. Quisque
sed dapibus nibh. Aenean non luctus leo. Phasellus luctus mauris id
aliquet dictum. Aliquam fermentum semper massa, vel dignissim nibh
dictum et. See Hubbard et al. (2021).

In luctus mauris vitae imperdiet luctus. Morbi volutpat ligula ut tortor
fermentum, eu ornare felis luctus. Donec semper diam vitae mattis
posuere. Suspendisse facilisis purus nisi, sit amet egestas ex tempor
ut. Cras tortor nulla, euismod at fermentum vel, dictum vel justo.
Aenean commodo interdum diam nec placerat. Nunc vestibulum felis at est
tincidunt, at euismod dui vestibulum. Nulla venenatis tortor at auctor
iaculis. Donec consectetur neque ut sagittis ornare. Nullam pharetra
felis tempor suscipit efficitur. Curabitur nibh ex, euismod at congue
hendrerit, egestas id mi. Duis porttitor neque in commodo elementum.
Fusce vitae fermentum nisi, euismod viverra augue. Curabitur at mi
pretium, accumsan purus nec, tempus turpis.

Donec non semper dui, quis aliquet est. Quisque quis sapien at massa
ultricies egestas. Duis consequat ultricies erat, a pulvinar nisl
vestibulum id. Sed tristique turpis ligula, et tempor lectus iaculis at.
Vivamus commodo sapien ac turpis vestibulum dapibus. Morbi tristique
arcu metus, et laoreet nisi varius nec. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Fusce
sit amet nisl at mauris suscipit aliquet. Nulla vitae dignissim urna.
Suspendisse sit amet arcu vitae magna blandit mattis. Vivamus convallis
efficitur pulvinar. Sed cursus elit nulla. Sed porta, arcu a euismod
pretium, odio dui lacinia lacus, ac vulputate nulla augue eget ex.
Nullam consequat ligula sit amet mattis aliquam. Nulla risus urna,
ultrices vel ullamcorper id, ornare viverra nunc.

Nunc in lobortis lacus. Duis maximus urna leo, varius sodales arcu
interdum nec. Pellentesque imperdiet dolor in leo eleifend dapibus. Ut
dapibus, lectus non viverra gravida, ipsum ex faucibus tellus, quis
iaculis risus tellus eget augue. Nullam a viverra est. Cras velit nisi,
interdum in lacus at, vehicula mattis elit. Curabitur eu viverra purus.
Proin pellentesque, metus vitae congue convallis, lorem metus feugiat
mi, sit amet auctor purus ligula bibendum ante. Nam id justo
scelerisque, rhoncus lectus in, fermentum libero. Donec tincidunt
egestas ex ac eleifend. Cras faucibus ipsum a nunc faucibus fermentum.
Integer et maximus lacus. Nam dictum nibh id viverra convallis.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\leavevmode\vadjust pre{\hypertarget{ref-Hubbard2021}{}}%
Hubbard, B., Christoffersen, P., Doyle, S. H., Chudley, T. R.,
Schoonman, C. M., Law, R., \& Bougamont, M. (2021). Borehole-based
characterization of deep mixed-mode crevasses at a greenlandic outlet
glacier. \emph{{AGU} Advances}, \emph{2}(2).
\url{https://doi.org/10.1029/2020av000291}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

River runoff is an important component of the global water cycle as it
comprises about one third of the precipitation over land areas
(Hagemann, 2020). Moreover, accurate runoff forecasting, especially over
extended periods, is pivotal for effective water resources management,
as highlighted by studies such as Yang et al.~(2018), Tan et al.~(2018),
and Fang et al.~(2019).

Over the past decades, models for long-term runoff forecasting have been
primarily bifurcated into physically based models and data-based models.
While the former attempts to emulate intricate and nonlinear physical
hydrological processes, the latter hinges on establishing statistical
models that delineate the relationship between large-scale climate
patterns and catchment runoff.

Machine Learning (ML) models, such as those employing artificial neural
networks, support vector machines, adaptive neuro-fuzzy inference
systems, and notably, Long Short-Term Memory (LSTM) neural networks,
have gained traction for long-term hydrological forecasting due to their
commendable performance (Humphrey et al 2016, Huang et al 2014, Ashrafi
et al 2017, Yuan et al 2018, Xu et al 2021).

LSTM networks, an evolution of the classical Recurrent Neural Networks
(RNNs), have shown stability and efficacy in sequence-to-sequence
predictions, such as using climatic indices for rainfall estimation or
long-term hydrological forecasting. However, a limitation of LSTMs is
their inability to effectively capture two-dimensional structures, an
area where Convolutional Neural Networks (CNNs) excel. Recognizing this,
we introduce the ConvLSTM, which integrates the strengths of both LSTM
and CNN, to extract spatiotemporal features from precipitation fields
for predicting river runoff in the Baltic Sea catchment, summarized by
97 inidivual rivers.

Modeling the Baltic Sea is to a large part the result of the quality of
the freshwater input, that is used for the simulation. Meier and
Kauker(2003) showed that decadal salinity variations of about 1 \(g\)
\(kg^{-1}\) are caused, inter alia, by annual runoff variations.
Further, Meier and Kauker (2003) showed that about 50 \% of the decadal
salinity variability can be explained by variations in freshwater input
into the Baltic Sea.

This paper delves into the application of deep learning, particularly
ConvLSTM, to the challenging task of precipitation nowcasting, a domain
yet to fully harness the potential of advanced machine learning
techniques. We present ConvLSTM as a novel solution to this
spatiotemporal sequence forecasting challenge, highlighting its
advantages and potential future applications.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{lstm-network}{%
\subsection{LSTM network}\label{lstm-network}}

The Long Short-Term Memory (LSTM), a specialized form of Recurrent
Neural Networks (RNNs), is specifically tailored for modeling temporal
sequences. Its unique design allows it to adeptly handle long-range
dependencies, setting it apart from traditional RNNs in terms of
accuracy. This prowess in modeling long-range dependencies has been
validated in various studies {[}12, 11, 17, 23{]}. The cornerstone of
LSTM's innovation is its memory cell, \(c_t\)t, which functions as a
repository of state information, also refered to as long-term memory.
This cell is accessed, modified, and reset through several
self-parameterized gates. As new inputs arrive, their information is
accumulated into the cell if the input gate is activated. The forget
gate \(f_t\) defines the percentage of the previous cell status that is
stored \(c_{t-1}\)\hspace{0pt}. The decision to propagate the latest
cell output, \(c_t\), to the final state, \(h_t\), is governed by the
output gate, \(o_t\). A significant advantage of this architecture is
the memory cell's ability to retain gradients. This mechanism addresses
the vanishing gradient problem, where, as input sequences elongate, the
influence of initial stages becomes harder to capture, causing gradients
of early input points to approach zero. The LSTM's activation function,
inherently recurrent, mirrors the identity function with a consistent
derivative of 1.0, ensuring the gradient remains stable throughout
backpropagation.

One LSTM cell hence maybe expressed as:

\[
\begin{aligned}
i_t &= \sigma(W_{xi} x_t + W_{hi} h_{t-1} + W_{ci} \circ c_{t-1} + b_i) \\
f_t &= \sigma(W_{xf} x_t + W_{hf} h_{t-1} + W_{cf} \circ c_{t-1} + b_f) \\
c_t &= f_t \circ c_{t-1} + i_t \circ \tanh(W_{xc} x_t + W_{hc} h_{t-1} + b_c) \\
o_t &= \sigma(W_{xo} x_t + W_{ho} h_{t-1} + W_{co} \circ c_t + b_o) \\
h_t &= o_t \circ \tanh(c_t)
\end{aligned}
\]

\hypertarget{convlstm-network}{%
\subsection{ConvLSTM network}\label{convlstm-network}}

The FC-LSTM, while adept in many scenarios, falters in encoding spatial
information when handling spatiotemporal data due to its reliance on
full connections in both input-to-state and state-to-state transitions.
Addressing this limitation, our design ensures that all inputs
\(X_1, \ldots, X_t\), cell outputs \(C_1, \ldots, C_t\), hidden states
\(H_1, \ldots, H_t\), and gates \(i_t, f_t, o_t\) of the ConvLSTM are 3D
tensors. The last two dimensions of these tensors represent spatial
dimensions, specifically rows and columns. Conceptually, these inputs
and states can be visualized as vectors positioned on a spatial grid.

In the ConvLSTM, the future state of a specific cell on this grid is
determined by the inputs and past states of its neighboring cells. This
spatial consideration is integrated by employing a convolution operator
in both state-to-state and input-to-state transitions, as illustrated in
Fig. 2. The foundational equations for ConvLSTM are:

\[
\begin{aligned}
i_t &= \sigma(W_{xi} \ast X_t + W_{hi} \ast H_{t-1} + W_{ci} \circ C_{t-1} + b_i) \\
f_t &= \sigma(W_{xf} \ast X_t + W_{hf} \ast H_{t-1} + W_{cf} \circ C_{t-1} + b_f) \\
C_t &= f_t \circ C_{t-1} + i_t \circ \tanh(W_{xc} \ast X_t + W_{hc} \ast H_{t-1} + b_c) \\
o_t &= \sigma(W_{xo} \ast X_t + W_{ho} \ast H_{t-1} + W_{co} \circ C_t + b_o) \\
H_t &= o_t \circ \tanh(C_t)
\end{aligned}
\]

Considering the states as representations of moving entities, a ConvLSTM
with a larger transitional kernel can capture rapid movements, while a
smaller kernel is apt for slower motions. Drawing parallels with
{[}16{]}, the traditional FC-LSTM's inputs, cell outputs, and hidden
states can also be envisioned as 3D tensors, with the last two
dimensions being unitary. Thus, the FC-LSTM can be seen as a specific
instance of ConvLSTM where all features are concentrated on a singular
cell.

To maintain consistency in the dimensions of states and inputs, padding
is essential before convolution operations. This padding, especially at
boundary points, can be interpreted as leveraging the state of external
surroundings for computations. Typically, all LSTM states are
initialized to zero before the first input, symbolizing a lack of
knowledge about future events. Similarly, zero-padding (as adopted in
this study) on hidden states sets the external world's state to zero,
implying no prior information about external conditions. This padding
approach allows for differential treatment of boundary points, proving
advantageous in many scenarios. For instance, in a system with a ball
moving and bouncing off unseen walls, the existence of these walls can
be inferred by observing the ball's repeated bounces, a deduction
challenging if boundary points share the same state transition dynamics
as inner points.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\vspace{1em}
